---
title: "Setting Up The Decision Space"
format: html
---

# Load Required Packages
```{r, warning=F, message=FALSE}
library(tidyverse)
library(here)
```

# Create a list with all combinations of the outcome variable
In the original paper by Kowall et al. (2025), the outcome variable was
composed of different concepts, which were composed of different variables 
in the data set. Therefore, I decided to first create a list of the four main
conceptual variables that were used in the original study. Then, I created a 
list with all possible combinations of those four conceptual variables.

```{r}
# First, create a list to conceptually combine the raw variables
outcome_options <- list(
  
  # "ever had heart attack or stroke"
  event_ever     = c("ph006d1",     # heart attack ever
                     "ph006d4"),    # stroke ever

  # "age at event"
  age_at_event   = c("ph009_1",     # age at heart attack
                     "ph009_4"),    # age at stroke

  # "event since last interview"
  since_last_int = c("ph072_1",     # heart attack since last interview
                     "ph072_2"),    # stroke since last interview

  # "death by event"
  death_by_event = c("xt011_")      # main cause of death: heart attack/stroke
)

# Then, create a list of all different combinations of the outcome options
  outcome_combinations <- unlist(

  # loop over combination sizes: 1 option, 2 options etc.
  lapply(1:length(outcome_options), function(k)

  # generate all combinations of group names of all possible sizes 
  combn(names(outcome_options), k, simplify = FALSE)
  ),

  # keep list-of-lists structure
  recursive = FALSE
)
  
## SELECT ONLY THE 7 THAT WERE ACTUALLY USED IN ORIGINAL MANY-ANALYST!

outcome_combinations <- c(outcome_combinations[15],
                          outcome_combinations[1],
                          outcome_combinations[5],
                          outcome_combinations[12],
                          outcome_combinations[11],
                          outcome_combinations[13],
                          outcome_combinations[7])
```

# Create a list with all combinations of the adjustment set
The adjustment set were made up of different variables. Because of time and  computational limitations, I decided to focus on those variables that were  chosen by at least four teams. To include these in my multiverse, I created a list with all possible combinations of covariates, including the empty set.

```{r}
# First, create a list of all adjustment options
adjustment_options = c("age_int",         # age at interview
                       "dn042_",          # sex
                       "country",         # country
                       "dn010_",          # education
                       "br001_",          # ever smoked daily
                       "phactiv")         # physical activity 

# Then, create a list of all different combinations of the adjustment options
adjustment_set <- unlist(
  
  # loop over combination sizes: 0 options (empty set), 1 option, 2 options etc.
  lapply(0:length(adjustment_options), function(k)
    
  # generate all combinations of group names of all possible sizes 
  combn(adjustment_options, k, simplify = FALSE)
  ),
  
  # keep list-of-lists structure
  recursive = FALSE
)
```

# Set Up The Decision Space
To create the decision space, I used `expand.grid` to create a data frame from all combinations of the decisions.

```{r}
decision_space_raw <- expand.grid(
 
  missing_data            = c("cc",              # complete cases
                              "ni"),             # "no info" category
  
  data_type               = c("long_w1",         # longitudinal from w1
                              "long_nostrict",   # longitudinal from w1-w6 
                              "cross_all",       # cross-sectional all waves
                              "cross_w1"),       # cross-sectional wave 1 only
  
  exclusion               = c("none",
                              "married_before18",     # year of marriage (before 18) - dn015_
                              "divorced_widowed",     # since when divorced (dn018_) / widowed (dn019_)
                              "country",              # (Israel)
                              "younger_than_55_w1",   # younger than 55 at w1 (dn003_)
                              "born_after1954"        # born after 1954 (dn003_)
                              ),
                            
  outcome                 = I(outcome_combinations),
  
  adjustment_set          = I(adjustment_set),   
                              
  
  model_type              = c("cox_haz",         # cox proportional hazards
                              "log_reg",         # logistic regression
                              "log-binom",       # log-binomial model
                              "mix_log_reg",     # mixed-effects logistic regression
                              "time_mix",        # discrete time mixed effect model
                              "GEE")             # generalized estimating equations
  )
```

This resulted in `r nrow(decision_space_raw)` different research paths.

```{r}
# Add columns for outcome and adjustment set where the combination of variables are + separated
# This is easier to read and use in later functions
decision_space <- decision_space_raw |>
  mutate(
    outcome_label = sapply(outcome, function(x) paste(sort(x), collapse = "+")),
    
    adjustment_label = sapply(adjustment_set, function(x) {
  if (length(x) == 0) "none" else paste(sort(x), collapse = "+")
})
  )
```

## Exclude invalid universes
```{r}
# Some models are not possible with specific data types
decision_space <- decision_space |>
  
  filter(!(data_type %in% c("cross_all", "cross_w1") & 
             model_type %in% c("cox_haz", "time_mix")) & # assumes longitudinal data
           
         !(data_type == "cross_w1" & 
             model_type %in% c("mix_log_reg", "GEE")) & # repeated measured needed
           
           !(data_type == "cross_w1" & grepl('since_last_int|death_by_event', outcome_label)) & # these variables do not exist in w1
           
        !(model_type == "cox_haz" & !(outcome_label %in% c("age_at_event+death_by_event+event_ever", "death_by_event+event_ever", "age_at_event+event_ever")))) # cox can only handle time-to-event variables so it needs age
```

# Write csv file to use the decision space dataset in the sampler
```{r}
write_csv(decision_space_raw, here("data/decision_space/decision_space_raw.csv"))
write_csv(decision_space, here("data/decision_space/decision_space.csv"))
```

