---
title: "Intro + Methods + Results Draft 1"
author: Loesje Ubbink
format:
  docx: default
---

```{r setup, echo = F, include = F, warning = F, message = F}
library(here)
library(tidyverse)
library(ggplot2)
library(dplyr)
library(tidyr)
library(kSamples)
library(tibble)
library(kableExtra)
library(stats)

decision_space_raw <- read_csv(here("data/decision_space/decision_space_raw.csv"))
decision_space <- read_csv(here("data/decision_space/decision_space.csv"))
sampled_universes <- readRDS(here("data/sampled_universes/sampled_universes_100.RDS"))
results <- readRDS(here("data/results/corrected_results_final_2026-04-28.RDS"))

# Make sure all NAs are seen as NAs
results <- results %>%
  mutate(RR = ifelse(model_type == "GEE" & (RR == Inf | RR == 0), NA, RR),
         conf.low = ifelse(model_type == "GEE" & (conf.low == Inf | conf.low == 0), NA, conf.low),
         conf.high = ifelse(model_type == "GEE" & (conf.high == Inf | conf.high == 0), NA, conf.high))

# How are NAs distributed per statistical model?
NAperModel <- results %>%
  group_by(model_type) %>%
  summarise(na_count = sum(is.na(RR)))

descRR <- results |>
  filter(!is.na(RR)) |>
  summarise(min = round(min(RR), 2),
    max = round(max(RR), 2),
    percSig = sum(p.value < 0.05)/nrow(results) * 100,
    median = round(median(RR), 2),
    mean = round(mean(RR), 2),
    sd = round(sd(RR), 2))

converged_results <- nrow(sampled_universes) - sum(is.na(results$RR))
```

# Introduction

Over recent years, it has become increasingly clear that the scientific community is facing a replication crisis [e.g., @forbes_supporting_2023]. A growing number of studies have reported difficulties reproducing previously published findings. This problem is often attributed to publication bias – the tendency to only publish significant results – and questionable research practices (QRPs). However, even studies aimed at removing the incentive for QRPs have painted a rather pessimistic picture of reproducible science [@auspurg_toward_2024]. Many-analyst studies, where multiple independent researchers answer the same research question using the same dataset, have shown considerable variation in analytic strategies and results [e.g., @silberzahn_many_2018; @bastiaansen_time_2020; @breznau_observing_2022; @hoogeveen_many-analysts_2023; @gould_same_2025; @kowall_marital_2025]. Reported effect sizes often vary from negative to positive, and from significant to non-significant, all within the same study. This means that there is not only considerable variation in the results but also in the substantive conclusions drawn from them. At first glance, this seems an immediate threat to the reproducibility of science. In an ideal world, one might expect that a single dataset should yield a single, 'true' answer, raising the question whether effect size variation should always be interpreted as problematic.

However, interpreting this variation without understanding its origins may not be that straightforward. A frequently raised criticism of many-analyst projects is that the posed research questions are often too vague, leaving analysts uncertain about how to proceed with data analysis and which variables should be included in their statistical models [@kummerfeld_one_2023; @hoogeveen_many-analysts_2023-1]. 


`For example, @silberzahn_many_2018 asked analysts to investigate the effect of skin tone of soccer players on the likelihood to receive a red card. This question can be interpreted different ways; are we interested in a descirptive association between skin tone and the likelihood of receiving a red card, or are we interested in a causal effect of skin tone on this likelihood? Depending on these different causal effects, different variables need to be included in the model.`

*researchers may choose to identify different causal effects, each of which has its merit, but each of which implies different estimands and corresponding identification strategies (reasonable counterfactuals; 15, 36). For example, one might be interested in (i) identifying the descriptive difference in the probability of receiving a red card by skin color, or one might be interested in (ii) identifying the causal direct effect of skin color (“racial bias”).*

This uncertainty is particularly problematic in situations where researchers are implicitly interested in finding causal relationships, which is often the case in many-analyst projects [@del_giudice_travelers_2021; @bulbulia_workflow_2023; @rohrer_what_2025]. In this context, including different confounders may correspond to different causal assumptions and thus different – potentially opposing – estimated effects. Therefore, when analysts select different variables, they may in fact be answering different questions [@simonsohn_specification_2015]. From this perspective, effect size variation may indicate meaningful substantive differences rather than problematic analytical noise. On the other hand, if the effect size variation stems from arbitrary, equally justifiable decisions whose options do not reflect different substantive meanings, this may be considered more problematic [@del_giudice_travelers_2021]. 

## The AB-framework
For better interpretation of effect size variation in many-analyst projects, @auspurg_toward_2024 have introduced a theoretical framework (hereafter referred to as the AB-framework) to distinguish between three different sources of variation, two of which are considered 'harmful' and one 'helpful' (see Table \ref{tab:AB_framework}).



First, effect size variation may arise from model uncertainty (a), for example in missing-data handling or modelling assumptions, reflecting ambiguity about the correct approach to estimate the parameter of interest and its underlying data-generating process. For example, if different missing data handling strategies (e.g., listwise deletion versus multiple imputation) result in different mean effect sizes, while the underlying observed data remain the same, this would imply that the outcome depends on an arbitrary analytic choice rather than on the data themselves. Secondly, effect size variation may be the result of sloppiness or hidden coding errors (b) made by the analysts. Both sources can be considered harmful, as they can introduce undesirable uncertainty into scientific analyses, making conclusions less credible.

In contrast, the third source concerns analytical choices that provide information about substantive variation that actually exists in the population (c). As mentioned earlier, selecting different target populations or confounders can lead to slightly different questions being answered. For example, if analyst X uses the full dataset, while analyst Y analyses only the men in that same dataset, X might find no effect in the general population, while Y finds that there is a male-specific positive effect. Given that the two analysts answered different questions, this source can be considered as helpful for the advancement of theories. By taking a closer look at the decisions made by analysts and the implications thereof, the AB-framework provides tools to interpret effect size variation in many-analyst projects.

## Identifying sources of variation in many-analyst projects
Although several many-analyst studies have aimed to identify the sources of effect size variation, so far no consensus has been reached [e.g.,@silberzahn_many_2018; @bastiaansen_time_2020; @breznau_observing_2022; @hoogeveen_many-analysts_2023; @gould_same_2025; @kowall_marital_2025]. For example, @hoogeveen_many-analysts_2023 found that the operationalisations of variables, rather than model choice,  influenced the results the most. In contrast, @gould_same_2025 showed that neither variation in variable selection nor model structure meaningfully predicted variability, but did not explicitly name another explanation source of variation. A similar pattern was found by @kowall_marital_2025, where data structure (cross-sectional versus longitudinal), rather than model or variable choice, likely had the largest impact on the results. Different studies thus highlight different sources of variation. 

Given the wide variety of topics, designs, and variables across projects, it may be unrealistic to expect a single, fixed set of explanatory factors that applies to all studies. However, even identifying a study-specific source of variation might not be possible based on many-analyst studies alone. A key limitation of these projects is that the involved analysts make decisions in an uncontrolled way [@kummerfeld_one_2023]. This leads to the problem of sparse data, where little data is generally available for each individual decision, as unique combinations of choices are often chosen only once. This sparsity makes it difficult to properly disentangle individual sources of variation and provide systematic quantitative evidence of their relative influence. Consequently, solely looking at analytical decisions in many-analyst projects may not provide enough information to interpret effect size variation with the AB-framework.

## Multiverse analysis
One approach to addressing the problem of data sparseness is multiverse analysis [e.g.,@steegen_increasing_2016; @hoogeveen_many-analysts_2023-1]. First, this method identifies all reasonable decisions, each with a specific number of options, and systematically combines them into a decision space [@liu_approximation_2023]. Since not all combinations may yield possible analyses, the set of valid analysis paths (universes) is always smaller than or equal to the full product. Then, results are reported across all valid universes. This way, multiverse analysis allows for the exploration of the entire decision space, rather than just a subset of choices made by analysts. As a result, substantially more data are available for each decision, enabling quantification of their specific impact on the results.

Although multiverse analysis seems a promising way of systematically evaluating the sensitivity of decisions, one of its limitations is that the decision space is often defined by a single researcher [@simonsohn_specification_2015]. Consequently, some reasonable analysis paths may be overlooked. However, by basing the multiverse analysis on an existing many-analyst project, this problem can be solved. The decision space will then be determined by the "wisdom of the crowd" and represent realistic research choices [@auspurg_toward_2024]. 

Another limitation is that different universes with contradicting substantial implications may be included within the same multiverse. @del_giudice_travelers_2021 argue that as a solution, a multiverse should only include truly arbitrary, but not theoretically meaningful decisions, as the latter may lead to misleading results. However, by including both types of decisions and subsequently classifying them with the AB-framework, results that might otherwise be considered misleading could instead be interpreted in a meaningful way.

## Present study
To assess the current state of the replication crisis using many-analyst projects, it is highly relevant to identify the most prominent sources of variation and their implications based on the AB-framework. Due to the descriptive nature of many-analyst studies, no consensus has yet been reached about what factors drive variability in the results, leading to the research question: 

*What is the relative contribution of harmful versus helpful analytical decisions in many-analyst projects to effect size variation when they are systematically evaluated in a multiverse?*

Aiming to gain quantitative evidence of the relative influence of different sources of variation, the present study will conduct a multiverse analysis based on the many-analyst project by @kowall_marital_2025, where 16 teams analysed whether marital status (cohabiting married versus never married) predicts cardiovascular disease. This was chosen as a case study, because there was a large variety of both harmful and helpful decisions, providing a rich decision space to evaluate the impact of a wide range of different decisions. The decision space consists of six analytical decisions: handling missing data, choice of exclusion criteria, statistical model and definitions of dependent variables, confounders and target population. Before running the analysis, these decisions are classified as harmful/helpful using the AB-framework. The multiverse analysis is largely based on sampling and approximation methods proposed by @liu_approximation_2023, where the decision space is sampled to immediately calculate the mean effect and decision sensitivity. Lastly, a specification curve will be plotted to visualise how the outcomes relate to different analysis paths. For more details, see the Methods section.

# Methods
## Ethics
Ethical approval was granted by the Ethics Review Board of the Faculty of Social and Behavioural Sciences of Utrecht University. 

## Data 
```{r dfs from all 4 DT sample size and number of part, echo=F, include=F, message=F, warning=F}
df_CA <- readRDS(here("data/DataTypeDfs/df_CA.RDS"))
df_CW1 <- readRDS(here("data/DataTypeDfs/df_CW1.RDS"))
df_LNS <- readRDS(here("data/DataTypeDfs/df_LNS.RDS"))
df_LW1 <- readRDS(here("data/DataTypeDfs/df_LW1.RDS"))
```

Consistent with @kowall_marital_2025, data from waves 1-7 (excluding wave 3) of the Survey of Health, Ageing and Retirement in Europe (SHARE) were used. SHARE is a longitudinal panel study conducted across 28 European countries, collecting data on health, social, economic and environmental policies, since 2004.

All variables needed for the construction of the decision space (outcome, exposure, adjustment, exclusion) were selected from different datasets that were available per wave. Then, one cleaned data set was created per wave, where the selected variables were linked through an ID variable. Lastly, all wave data sets were combined based on the selected Data Type option in the multiverse analysis. After data cleaning, but before applying the exclusion criteria and missing data handling, the samples consisted of:

- cross-secional all waves: `r format(length(unique(df_CA$mergeid)), big.mark = ",")` participants across `r format(nrow(df_CA), big.mark = ",")` observations.
- cross-sectional wave 1: `r format(length(unique(df_CW1$mergeid)), big.mark = ",")` participants across `r format(nrow(df_CW1), big.mark = ",")` observations.
- longitudinal no strict baseline: `r format(length(unique(df_LNS$mergeid)), big.mark = ",")` participants across `r format(nrow(df_LNS), big.mark = ",")` observations.
- longitudinal wave 1: `r format(length(unique(df_LW1$mergeid)), big.mark = ",")` participants across `r format(nrow(df_LW1), big.mark = ",")` observations.

## Definition of decision space 
To define the decision space, all options of six different decisions made by the teams in @kowall_marital_2025 were identified based on information provided in the original publication (see Table \ref{tab:decision_space}). When systematically combining all options across decisions, the decision space consisted of `r format(nrow(decision_space_raw), big.mark = ",")` universes. After removing invalid research paths, `r format(nrow(decision_space), big.mark = ",")` valid universes remained. Invalid paths included combinations of cross-sectional data with models requiring longitudinal data structure (Cox proportional hazards (Cox PH), discrete-time mixed-effects), cross-sectional wave 1 data with models requiring repeated measures (mixed-effects logistic regression, generalised estimating equations (GEE)), wave 1 data with outcomes that were not available at baseline (since last interview, cause of death), and Cox models with outcomes lacking time-to-event variables.

## Classifying sources of variation using the AB-framework
The selected decisions were classified using the AB-framework. Missing data and statistical model choice were classified as harmful sources of variation. Operationalisation of the outcome variable, selection of confounders, and exclusion criteria were classified as helpful. Data structure was also classified as helpful, as the framework treats variations over time as representing different target populations (see Table \ref{tab:decision_space}).


## Converting OR and HR to RR
The statistical models yielded different effect size measures. Logistic regression, mixed-effects logistic regression, GEE, and discrete-time mixed-effects models produced odds ratios (OR). Cox PH models produced hazard ratios (HR). Log-binomial models produced risk ratios (RR) directly. To enable comparison across universes, all effect sizes were converted to RRs. Odds ratios and hazard ratios both approximate risk ratios when the outcome is rare, but can diverge substantially when the outcome is common [@vanderweele_optimal_2020]. Given that the overall prevalence in the unexposed group ($p_0$) was around 18% in all subsets of the data, conversions for both OR and HR were necessary.

Conversions to risk ratios often require knowing $p_0$ [@zhang_whats_1998]. However, in a multiverse analysis, $p_0$ may vary across universes, depending on data type and exclusion decisions resulting in different data subsets. This could introduce arbitrary variation in the results, distorting the actual decision sensitivity.

Therefore, conversions proposed by @vanderweele_optimal_2020 were used, designed for scenarios when $p_0$ is unknown but exceeds 10%. For odds ratios [@huang_development_2022]:

\begin{align}
  RR \approx \sqrt{OR}
\end{align}

For hazard ratios, where prevalence is estimated between 20% and 80% [@vanderweele_optimal_2020; @huang_development_2022-1]:

\begin{align}
  RR \approx \frac{1 - 0.5^{\sqrt{HR}}}{1 - 0.5^{\sqrt{1 / HR}}}
\end{align}

Although the overall prevalence was slightly lower than 20%, a conversion not requiring a specific $p_0$ value was preferred to avoid introducing additional variation.

## Round Robin sampling
Because the number of reasonable universes grows exponentially with the number of decisions and their options, the computational cost of multiverse analyses increases rapidly. To address this, @liu_approximation_2023 proposed sampling universes rather than exploring the full decision space. They evaluated various algorithms and found that Round Robin sampling was one of the most effective methods for approximating multiverse results. Therefore, this method was applied in the present study. Round Robin sampling aims to represent each decision option at least once. The algorithm iteratively loops through all decisions and their options, sampling one universe at each stage without replacement, ensuring each sampled universe is included at most once. For a visual example of how the algorithm works, see @fig-roundrobin. 

Although @liu_approximation_2023 proposed sampling universes one by one and immediately processing the multiverse results, in this study universes were selected *a priori* before running the multiverse analysis. The sampler was implemented in two steps. First, decision options were mapped to corresponding row numbers in the full decision space. Then, row numbers were sampled according to the Round Robin algorithm and used to select universes from the original decision space. This approach was more computationally efficient than sampling directly from the decision space. A random seed was set to ensure reproducibility.

To ensure sufficient data were available for each option to quantify decision sensitivity, a minimum of 100 universes per option was sampled. As adjustment set had the most options (64), this resulted in `r format(nrow(sampled_universes), big.mark = ",")` sampled universes in total.


## Decision sensitivity
After converting all effect size estimates to risk ratios, density plots were used to visualise RR distributions across decision options. In addition, to visually assess the robustness of multiverse outcomes, each universe was plotted against its estimated effect size in a specification curve [@simonsohn_specification_2015]. However, as both methods only provide descriptive information about the distribution of results rather than quantifying decision sensitivity, they were used as visual aids to inspect whether the outcome distribution was robust to different analytical specifications.

To quantify the decision sensitivity - the extent to which different options within each analytical decision lead to different outcomes - k-sample Anderson-Darling tests (AD tests) were used, as proposed by @liu_approximation_2023. This test evaluates whether the RRs resulting from *k* options within a decision stem from the same empirical cumulative distribution function (ECDF). If this null hypothesis was rejected, the decision was classified as sensitive. The AD statistic measures the weighted difference between the ECDFs of *k* samples. The standardised AD statistic (T.AD), which adjusts for number of groups and sample size, was used for between-decision comparisons. Higher T.AD values indicate more sensitive decisions.

Because there were 64 different adjustment sets, Kolmogorov-Smirnov tests (KS tests) were used for a more detailed evaluation of its decision sensitivity, comparing RR distributions for specifications including versus excluding each adjustment variable. The KS test statistic (D) represents the maximum absolute difference between the empirical cumulative distribution functions of the two samples, ranging from 0 (identical) to 1 (no overlap). Values closer to 0 indicate a higher likelihood that the two samples were drawn from the same distribution.

## Supplementary Materials
All analyses were conducted in R [version 4.4.2, @r_core_team_r_2024]. All code scripts containing sampler and multiverse functions, and package versions documented in the lock file, can be found [here](https://github.com/lodewijn/ManyAnalystThesisCompendium).

# Results

## Model convergence
Out of the `r format(nrow(sampled_universes), big.mark = ",")` sampled universes, `r format(sum(is.na(results$RR)), big.mark = ",")` failed to converge. Non-convergence mostly occurred in log-binomial models (n = `r format(NAperModel$na_count[NAperModel$model_type == "log-binom"], big.mark = ",")`), followed by generalised estimating equations (GEE; n = `r format(NAperModel$na_count[NAperModel$model_type == "GEE"], big.mark = ",")`) and logistic regression (n = `r format(NAperModel$na_count[NAperModel$model_type == "log_reg"], big.mark = ",")`). The other three model types (Cox PH, mixed effects logistic, and discrete-time mixed effects) converged successfully in all cases. Across other analytical decisions, non-convergence was distributed approximately equally, suggesting these decisions did not substantially affect model convergence. In total, `r format(converged_results, big.mark = ",")` universes yielded valid risk ratio estimates. The number of converged and non-converged universes per decision option can be found in @tbl-RRfulltable.

## Overall multiverse results
Across the `r format(converged_results, big.mark = ",")` successful specifications, risk ratio estimates ranged from `r descRR$min` to `r descRR$max` (median = `r descRR$median`, mean = `r descRR$mean`, SD = `r descRR$sd`), with `r round(descRR$percSig, 1)`% of the universes yielding significant results (p < .05). Within this sample of the full multiverse, marriage thus appeared both as a protective factor (RR < 1) and a risk factor (RR > 1) for cardiovascular events, depending on the analytical specification. @fig-spec-curve shows the full distribution of the effect estimates across universes, arranged from smallest to largest risk ratio, with significance indicated by colour.

Several patterns emerged in the specification curve (@fig-spec-curve). Universes including age as an adjustment variable were clustered around higher RRs (mostly $\geq 1$), while universes without age were clustered around lower RRs (all < 1), suggesting a potential effect of age on the relationship between marriage and cardiovascular events. The highest RRs (> 1.25) only occurred in specifications including Cox PH models, although almost none of these estimates reached significance. Log-binomial models, in contrast, yielded only RRs below 1, almost all of which were significant. Furthermore, effect estimates seemed robust to data type (except cross-sectional W1), missing data handling, outcomes (except Age + Ever), and exclusion criteria, with no clear clustering patterns across these decisions.

## Decision sensitivity
```{r ad setup, echo = F, include = F, warning = F, message = F}
set.seed(260722)
DT_AD <- ad.test(RR ~ data_type, data = results)
MD_AD <- ad.test(RR ~ missing_data, data = results)
EC_AD <- ad.test(RR ~ exclusion, data = results)
OUT_AD <- ad.test(RR ~ outcome_label, data = results)
ADJ_AD <- ad.test(RR ~ adjustment_label, data = results)
SM_AD <- ad.test(RR ~ model_type, data = results)

ADtable <- tibble(
  Decision = c("Data Type", "Missing Data", "Exclusion Criteria", 
               "Outcome", "Statistical Model", "Adjustment Set"), 
  `N options` = c(DT_AD$k, MD_AD$k, EC_AD$k, OUT_AD$k, SM_AD$k, ADJ_AD$k),
  T.AD = c(DT_AD$ad[1,2], MD_AD$ad[1,2], EC_AD$ad[1,2], 
             OUT_AD$ad[1,2], SM_AD$ad[1,2], ADJ_AD$ad[1,2]),
  p_value = c(DT_AD$ad[1,3], MD_AD$ad[1,3], EC_AD$ad[1,3], 
                OUT_AD$ad[1,3], SM_AD$ad[1,3], ADJ_AD$ad[1,3])
) |>
  arrange(T.AD)
```
```{r full RR table setup, echo = F, include = F, warning = F, message =F}
decisions <- c("model_type", "data_type", "missing_data", "exclusion", "outcome_label", "adjustment_label")

# Create label mappings
decision_labels <- c(
  "data_type" = "Data Type",
  "missing_data" = "Missing Data",
  "exclusion" = "Exclusion Criteria",
  "outcome_label" = "Outcome",
  "model_type" = "Statistical Model",
  "adjustment_label" = "Adjustment Set"
)

option_labels <- c(
  # Data type
  "cross_all" = "Cross-sectional (all waves)",
  "cross_w1" = "Cross-sectional (wave 1)",
  "long_nostrict" = "Longitudinal (no strict baseline)",
  "long_w1" = "Longitudinal (wave 1)",
  
  # Missing data
  "cc" = "Complete cases",
  "ni" = "No information",
  
  # Exclusion
  "born_after1954" = "Born after 1954",
  "country" = "Country",
  "divorced_widowed" = "Divorced/widowed",
  "married_before18" = "Married before 18",
  "none" = "None",
  "younger_than_55_w1" = "Younger than 55",
  
  # Model type
  "cox_haz" = "Cox PH",
  "log_reg" = "Logistic regression",
  "log-binom" = "Log-binomial",
  "mix_log_reg" = "Mixed effects logistic regression",
  "time_mix" = "Discrete-time mixed effects",
  "GEE" = "GEE"
)

# Create table with nice labels
full_table <- map_dfr(decisions, function(d) {
  results %>%
    group_by(Option = .data[[d]]) %>%
    summarise(
      N_total = n(),
      N_failed = sum(is.na(RR)),
      N_valid = sum(!is.na(RR)),
      `Mean RR` = round(mean(RR, na.rm = TRUE), 2),
      `RR Range` = paste0("[", round(min(RR, na.rm = TRUE), 2), ", ", round(max(RR, na.rm = TRUE), 2), "]"),
      `% Sig` = paste0(round(mean(p.value < 0.05, na.rm = TRUE) * 100, 0), "%"),
      .groups = "drop"
    ) %>%
    arrange(`Mean RR`) %>%
    mutate(
      Decision = decision_labels[d],
      Option = ifelse(Option %in% names(option_labels), option_labels[Option], Option)
    ) %>%
    select(Decision, Option, N_valid, N_failed, `Mean RR`, `RR Range`, `% Sig`)
})

# Clean adjustment labels in the data first
full_table <- full_table %>%
  mutate(
    Option = Option %>%
      str_replace_all("dn010_", "Edu") %>%
      str_replace_all("dn042_", "Sex") %>%
      str_replace_all("br001_", "Smoke") %>%
      str_replace_all("phactiv", "Phys") %>%
      str_replace_all("country", "Country") %>%
      str_replace_all("age_int", "Age") %>%
      str_replace_all("none", "None") %>%
      str_replace_all("since_last_int", "Since") |>
      str_replace_all("age_at_event", "Age event") %>%
      str_replace_all("death_by_event", "Death") %>%
      str_replace_all("event_ever", "Event ever") %>%
      str_replace_all("\\+", " + ")
  )

summary_full_table <- full_table %>%
    group_by(Decision) %>%
    summarise(
      min = round(min(`Mean RR`), 2),
      max = round(max(`Mean RR`), 2),   
      range = abs(round(min(`Mean RR`) - max(`Mean RR`),2)),
      .groups = "drop"
    ) |>
  arrange(range)
```
```{r KS test setup, echo = F, include = F, warning = F, message = F}
df_adj <- results %>%
  filter(!is.na(RR)) |>
  mutate(
    has_none = grepl("none", adjustment_label, ignore.case = TRUE),
    has_age = grepl("age_int", adjustment_label, ignore.case = TRUE),
    has_sex = grepl("dn042_", adjustment_label, ignore.case = TRUE),
    has_country = grepl("country", adjustment_label, ignore.case = TRUE),
    has_smoking = grepl("br001_", adjustment_label, ignore.case = TRUE),
    has_education = grepl("dn010_", adjustment_label, ignore.case = TRUE),
    has_activity = grepl("phactiv", adjustment_label, ignore.case = TRUE)
  )

# KS test for each confounder
ks_none <- ks.test(df_adj$RR[df_adj$has_none], df_adj$RR[!df_adj$has_none])
ks_age <- ks.test(df_adj$RR[df_adj$has_age], df_adj$RR[!df_adj$has_age])
ks_sex <- ks.test(df_adj$RR[df_adj$has_sex], df_adj$RR[!df_adj$has_sex])
ks_country <- ks.test(df_adj$RR[df_adj$has_country], df_adj$RR[!df_adj$has_country])
ks_smoking <- ks.test(df_adj$RR[df_adj$has_smoking], df_adj$RR[!df_adj$has_smoking])
ks_education <- ks.test(df_adj$RR[df_adj$has_education], df_adj$RR[!df_adj$has_education])
ks_activity <- ks.test(df_adj$RR[df_adj$has_activity], df_adj$RR[!df_adj$has_activity])

# Combine results
ks_results <- data.frame(
  `Adjustment variable` = c("None", "Age", "Sex", "Country", "Smoking", "Education", "Physical activity"),
  Included = c(sum(df_adj$has_none), sum(df_adj$has_age), sum(df_adj$has_sex), 
               sum(df_adj$has_country), sum(df_adj$has_smoking), sum(df_adj$has_education), 
               sum(df_adj$has_activity)),
  Excluded = c(sum(!df_adj$has_none), sum(!df_adj$has_age), sum(!df_adj$has_sex), 
               sum(!df_adj$has_country), sum(!df_adj$has_smoking), sum(!df_adj$has_education), 
               sum(!df_adj$has_activity)),
  D = c(ks_none$statistic, ks_age$statistic, ks_sex$statistic, ks_country$statistic, 
        ks_smoking$statistic, ks_education$statistic, ks_activity$statistic),
  p.value = c(ks_none$p.value, ks_age$p.value, ks_sex$p.value, ks_country$p.value,
        ks_smoking$p.value, ks_education$p.value, ks_activity$p.value), 
  check.names = F
)

ks_results$D <- round(ks_results$D, 3)
ks_results$p.value <- ifelse(ks_results$p.value < 0.001, "< .001", round(ks_results$p.value, 3))

ks_results <- ks_results |>
  arrange(D)
```
```{r age and sex across groups, echo = F, include = F, warning = F, message = F}
df <- readRDS(here("data/cleandata/df_all_waves.RDS"))
source(here("scripts/04_set_up_multiverse/03_exposure_function.R"))
df <- exposure_function(df)

descSHARE <- df |>
  group_by(exposure) |>
  summarise(meanAGE = mean(age_int, na.rm=T), 
            NMale = sum(ifelse(dn042_ == "Male", 1, 0)),
            NFemale = sum(ifelse(dn042_ == "Female", 1, 0)), 
            pctFemale = mean(dn042_ == "Female", na.rm = TRUE) * 100)
```

To examine the effect size distributions across decision options in more detail, density plots were used (see @fig-dens-plot). Data type, missing data, exclusion criteria, and outcome showed largely overlapping RR distributions across their options. However, in case of statistical models and adjustment sets there were clear separations in RR distributions between options. Consistent with the pattern shown in @fig-spec-curve, log-binomial models were clustered below RR values of 1, while the density of Cox PH showed a long tail extending above 1. Similarly, specifications including age in the adjustment set showed a clear shift towards higher RRs compared to those without age as a confounder.

Mean RR ranges across options confirmed the patterns found in the specification curve and density plots (@tbl-RRfulltable). Statistical model (`r summary_full_table$min[6]` - `r summary_full_table$max[6]`) and adjustment set (`r summary_full_table$min[5]` - `r summary_full_table$max[5]`) showed the largest variation in mean RR across options, while other decisions showed smaller ranges: exclusion criteria (`r summary_full_table$min[1]` - `r summary_full_table$max[1]`), missing data (`r summary_full_table$min[2]` - `r summary_full_table$max[2]`), data type (`r summary_full_table$min[3]` - `r summary_full_table$max[3]`), and outcome (`r summary_full_table$min[4]` - `r summary_full_table$max[4]`).

A k-sample Anderson-Darling test (AD test) was used to quantify decision sensitivity (for the ECDF per decision see @fig-ECDF). As shown in @tbl-ADtest, AD tests for all decisions were highly significant (almost all p < .001), indicating distributional differences across options within each decision. Adjustment set and statistical model showed substantially higher T.AD values (`r round(ADtable$T.AD[5], 2)` and `r round(ADtable$T.AD[6],2)`, respectively) than other decisions (missing data: `r round(ADtable$T.AD[1], 2)`; exclusion: `r round(ADtable$T.AD[2], 2)`; outcome: `r round(ADtable$T.AD[3], 2)`; data type: `r round(ADtable$T.AD[4], 2)`). This indicates that statistical model and adjustment set had the largest impact on the distribution of effect estimates. The number of options per decision did not seem to substantially affect the T.AD values.



Kolmogorov-Smirnov tests (KS test) were used to compare RR distributions for specifications including versus excluding each adjustment variable. For the ECDFs across adjustment variables, see @fig-ECDF-ADJ. As can be seen in @tbl-KStest, inclusion of all adjustment variables, except smoking, resulted in significantly different distributions than when these variables were excluded (p < .001). Age showed the largest effect (`r round(ks_results$D[7],2)`), followed by not including any confounders (`r round(ks_results$D[6],2)`) and sex (`r round(ks_results$D[5],2)`). Physical activity, country, and education showed moderate, but significant, effects ($D_{activity}$ = `r round(ks_results$D[4],2)`, $D_{country}$ = `r round(ks_results$D[3],2)`, and $D_{education}$ = `r round(ks_results$D[2],2)`). This confirms that the adjustment set sensitivity seems to be driven primarily by age, sex and the decision not to include confounders at all.


The adjustment set sensitivity driven by age, sex and none is also visible in the RR density distributions (@fig-dens-ADJ). Including age as an adjustment variable shifted the RR distribution to the right (higher RRs), while including sex shifted it to the left (lower RRs). Specifications without any confounders also showed lower RRs. Other adjustment variables did not result in pronounced differences between RR distributions.

In the underlying data, married individuals were on average `r abs(round(descSHARE$meanAGE[descSHARE$exposure == 1] - descSHARE$meanAGE[descSHARE$exposure == 0],0))` years younger than unmarried individuals (`r round(descSHARE$meanAGE[descSHARE$exposure == 1],0)` versus `r round(descSHARE$meanAGE[descSHARE$exposure == 0],0)` years old). Additionally, the distribution of sex differed between groups: `r round(descSHARE$pctFemale[descSHARE$exposure == 1], 0)`% of married individuals were female, compared to `r round(descSHARE$pctFemale[descSHARE$exposure == 0], 0)`% in the unmarried group.


# Discussion



\newpage

# References

::: {#refs}
:::

\newpage

# Appendix {#sec-appendix .appendix}




