---
title: "Results Draft 2 Word"
author: Loesje Ubbink
format:
  docx: default
---

```{r setup, echo = F, include = F, warning = F, message = F}
library(here)
library(tidyverse)
library(ggplot2)
library(dplyr)
library(tidyr)
library(kSamples)
library(tibble)
library(kableExtra)

decision_space_raw <- read_csv(here("data/decision_space/decision_space_raw.csv"))
decision_space <- read_csv(here("data/decision_space/decision_space.csv"))
sampled_universes <- readRDS(here("data/sampled_universes/sampled_universes_100.RDS"))
results <- readRDS(here("data/results/all_models_final_results_Univ100_6400_2026-04-20.RDS"))

# Make sure all NAs are seen as NAs
results <- results %>%
  mutate(RR = ifelse(model_type == "GEE" & (RR == Inf | RR == 0), NA, RR),
         conf.low = ifelse(model_type == "GEE" & (conf.low == Inf | conf.low == 0), NA, conf.low),
         conf.high = ifelse(model_type == "GEE" & (conf.high == Inf | conf.high == 0), NA, conf.high))

# How are NAs distributed per statistical model?
NAperModel <- results %>%
  group_by(model_type) %>%
  summarise(na_count = sum(is.na(RR)))

descRR <- results |>
  filter(!is.na(RR)) |>
  summarise(min = round(min(RR), 2),
    max = round(max(RR), 2),
    percSig = sum(p.value < 0.05)/nrow(results) * 100,
    median = round(median(RR), 2),
    mean = round(mean(RR), 2),
    sd = round(sd(RR), 2))

converged_results <- nrow(sampled_universes) - sum(is.na(results$RR))
```

# Results

## Model convergence
Out of the `r format(nrow(sampled_universes), big.mark = ",")` sampled universes, `r format(sum(is.na(results$RR)), big.mark = ",")` failed to converge. Non-convergence mostly occurred in log-binomial models (n = `r format(NAperModel$na_count[NAperModel$model_type == "log-binom"], big.mark = ",")`), followed by generalised estimating equations (GEE; n = `r format(NAperModel$na_count[NAperModel$model_type == "GEE"], big.mark = ",")`) and logistic regression (n = `r format(NAperModel$na_count[NAperModel$model_type == "log_reg"], big.mark = ",")`). The other three model types (Cox PH, mixed effects logistic, and discrete-time mixed effects) converged successfully in all cases. Across other analytical decisions, non-convergence was distributed approximately equally, suggesting these decisions did not substantially affect model convergence. In total, `r format(converged_results, big.mark = ",")` universes yielded valid risk ratio estimates. The number of converged and non-converged universes per decision option can be found in @tbl-RRfulltable.

## Overall multiverse results
Across the `r format(converged_results, big.mark = ",")` successful specifications, risk ratio estimates ranged from `r descRR$min` to `r descRR$max` (median = `r descRR$median`, mean = `r descRR$mean`, SD = `r descRR$sd`), with `r round(descRR$percSig, 1)`% of the universes yielding significant results (p < .05). Within this sample of the full multiverse, marriage thus appeared both as a protective factor (RR < 1) and a risk factor (RR > 1) for cardiovascular events, depending on the analytical specification. @fig-spec-curve shows the full distribution of the effect estimates across universes, arranged from smallest to largest risk ratio, with significance indicated by colour.

Several patterns emerged in the specification curve (@fig-spec-curve). Universes including age as an adjustment variable were clustered around higher RRs (mostly $\geq 1$), while universes without age were clustered around lower RRs (all < 1), suggesting a potential effect of age on the relationship between marriage and cardiovascular events. The highest RRs (> 1.2) only occurred in specifications including Cox PH models, although almost none of these estimates reached significance. Log-binomial models, in contrast, yielded only RRs below 1, almost all of which were significant. Furthermore, effect estimates seemed robust to data type (except cross-sectional W1), missing data handling, outcomes (except Age + Ever), and exclusion criteria, with no clear clustering patterns across these decisions.



## Decision sensitivity
```{r ad setup, echo = F, include = F, warning = F, message = F}
set.seed(260722)
DT_AD <- ad.test(RR ~ data_type, data = results)
MD_AD <- ad.test(RR ~ missing_data, data = results)
EC_AD <- ad.test(RR ~ exclusion, data = results)
OUT_AD <- ad.test(RR ~ outcome_label, data = results)
ADJ_AD <- ad.test(RR ~ adjustment_label, data = results)
SM_AD <- ad.test(RR ~ model_type, data = results)

ADtable <- tibble(
  Decision = c("Data Type", "Missing Data", "Exclusion Criteria", 
               "Outcome", "Statistical Model", "Adjustment Set"), 
  `N options` = c(DT_AD$k, MD_AD$k, EC_AD$k, OUT_AD$k, SM_AD$k, ADJ_AD$k),
  T.AD = c(DT_AD$ad[1,2], MD_AD$ad[1,2], EC_AD$ad[1,2], 
             OUT_AD$ad[1,2], SM_AD$ad[1,2], ADJ_AD$ad[1,2]),
  p_value = c(DT_AD$ad[1,3], MD_AD$ad[1,3], EC_AD$ad[1,3], 
                OUT_AD$ad[1,3], SM_AD$ad[1,3], ADJ_AD$ad[1,3])
) |>
  arrange(T.AD)
```
```{r full RR table setup, echo = F, include = F, warning = F, message =F}
decisions <- c("model_type", "data_type", "missing_data", "exclusion", "outcome_label", "adjustment_label")

# Create label mappings
decision_labels <- c(
  "data_type" = "Data Type",
  "missing_data" = "Missing Data",
  "exclusion" = "Exclusion Criteria",
  "outcome_label" = "Outcome",
  "model_type" = "Statistical Model",
  "adjustment_label" = "Adjustment Set"
)

option_labels <- c(
  # Data type
  "cross_all" = "Cross-sectional (all waves)",
  "cross_w1" = "Cross-sectional (wave 1)",
  "long_nostrict" = "Longitudinal (no strict baseline)",
  "long_w1" = "Longitudinal (wave 1)",
  
  # Missing data
  "cc" = "Complete cases",
  "ni" = "No information",
  
  # Exclusion
  "born_after1954" = "Born after 1954",
  "country" = "Country",
  "divorced_widowed" = "Divorced/widowed",
  "married_before18" = "Married before 18",
  "none" = "None",
  "younger_than_55_w1" = "Younger than 55",
  
  # Model type
  "cox_haz" = "Cox PH",
  "log_reg" = "Logistic regression",
  "log-binom" = "Log-binomial",
  "mix_log_reg" = "Mixed effects logistic regression",
  "time_mix" = "Discrete-time mixed effects",
  "GEE" = "GEE"
)

# Create table with nice labels
full_table <- map_dfr(decisions, function(d) {
  results %>%
    group_by(Option = .data[[d]]) %>%
    summarise(
      N_total = n(),
      N_failed = sum(is.na(RR)),
      N_valid = sum(!is.na(RR)),
      `Mean RR` = round(mean(RR, na.rm = TRUE), 2),
      `RR Range` = paste0("[", round(min(RR, na.rm = TRUE), 2), ", ", round(max(RR, na.rm = TRUE), 2), "]"),
      `% Sig` = paste0(round(mean(p.value < 0.05, na.rm = TRUE) * 100, 0), "%"),
      .groups = "drop"
    ) %>%
    arrange(`Mean RR`) %>%
    mutate(
      Decision = decision_labels[d],
      Option = ifelse(Option %in% names(option_labels), option_labels[Option], Option)
    ) %>%
    select(Decision, Option, N_valid, N_failed, `Mean RR`, `RR Range`, `% Sig`)
})

# Clean adjustment labels in the data first
full_table <- full_table %>%
  mutate(
    Option = Option %>%
      str_replace_all("dn010_", "Edu") %>%
      str_replace_all("dn042_", "Sex") %>%
      str_replace_all("br001_", "Smoke") %>%
      str_replace_all("phactiv", "Phys") %>%
      str_replace_all("country", "Country") %>%
      str_replace_all("age_int", "Age") %>%
      str_replace_all("none", "None") %>%
      str_replace_all("since_last_int", "Since") |>
      str_replace_all("age_at_event", "Age event") %>%
      str_replace_all("death_by_event", "Death") %>%
      str_replace_all("event_ever", "Event ever") %>%
      str_replace_all("\\+", " + ")
  )

summary_full_table <- full_table %>%
    group_by(Decision) %>%
    summarise(
      min = round(min(`Mean RR`), 2),
      max = round(max(`Mean RR`), 2),   
      range = abs(round(min(`Mean RR`) - max(`Mean RR`),2)),
      .groups = "drop"
    ) |>
  arrange(range)
```

To get a more detailed overview of the effect size distributions across decision options, density plots were used (see @fig-dens-plot). Data type, missing data, exclusion criteria, and outcome showed largely overlapping RR distributions across their options. However, in case of statistical models and adjustment sets there were clear separations in RR distributions between options. Consistent with the pattern shown in @fig-spec-curve, log-binomial models were clustered below RR values of 1, while the densitiy of Cox PH showed a large tail with value well above 1. Similarly, specifications including age as in the adjustment set showed a clear shift towards lower RR values, compared to those without age as a confounder.

Mean RR ranges across options confirmed the patterns found in the specification curve and density plots (@tbl-RRfulltable). Statistical model (`r summary_full_table$min[6]` - `r summary_full_table$max[6]`) and adjustment set (`r summary_full_table$min[5]` - `r summary_full_table$max[5]`) showed the largest variation in mean RR across options, while other decisions showed smaller ranges: exclusion criteria (`r summary_full_table$min[1]` - `r summary_full_table$max[1]`), missing data (`r summary_full_table$min[2]` - `r summary_full_table$max[2]`), data type (`r summary_full_table$min[3]` - `r summary_full_table$max[3]`), outcome (`r summary_full_table$min[4]` - `r summary_full_table$max[4]`).


When looking at the ranges of mean RRs per option (see @tbl-RRfulltable for the exact values per option) same pattern emerges, with statistical model and adjustment set showing the largest variation in mean RR across options ([`r summary_full_table$min[6]` - `r summary_full_table$max[6]`] and [`r summary_full_table$min[5]` - `r summary_full_table$max[5]`], respectively), while the mean ranges (min - max) in the other decisions were all smaller ($\leq 0.08$).

*"Mean RRs by statistical model ranged from 0.83 (log-binomial) to 1.12 (Cox PH). For adjustment sets, specifications excluding age yielded mean RRs of 0.83–0.91, while those including age yielded 0.97–1.07. Other decisions showed narrower ranges: data structure (0.93–0.97), missing data (0.94–0.95), exclusion criteria (0.94–0.95), and outcome (0.92–1.00). Full results by option are provided in @tbl-full."*


In addition to visually describing patterns in result across universes, a k-samples Anderson-Darling test (AD-test) was used to quantify the decision sensitivity, testing the hypothesis that the results from k individual samples (options) all come from the same empirical distribution function. As shown in @tbl-ADtest, AD tests for all decisions were highly significant (almost all p < .001), indicating distributional differences across options within each decision. The standardised AD statistic (T.AD), which corrects for number of groups and sample size, was used for between-decision comparisons. Adjustment set and statistical model showed substantially higher T.AD values (`r round(ADtable$T.AD[5], 2)` - `r round(ADtable$T.AD[6],2)`) than other decisions (`r round(ADtable$T.AD[1],2)` - `r round(ADtable$T.AD[4], 2)`), indicating these decisions had the largest impact on the distribution of effect estimates. 



# Appendix {#sec-appendix .appendix}

